\chapter*{About the Author}
\addcontentsline{toc}{chapter}{About the Author}
\markboth{About the Author}{About the Author}

\begin{center}
{\color{oreillyteal}\rule{0.3\textwidth}{2pt}}
\end{center}

\vspace{1em}

\noindent\textbf{Sanchit Alekh} is a Customer Engineer for AI and Machine Learning at Google, based in Munich, Germany. He works with enterprise clients across the DACH region to conceptualize, architect, and deploy AI/ML solutions on Google Cloud Platform---from foundation model integration and RAG pipelines to large-scale inference infrastructure.

Before joining Google, Sanchit served as a Cloud Solution Architect at Microsoft, where he designed and implemented cloud-native AI systems for enterprise customers across Europe. His career spans the full arc of the modern AI revolution: from classical machine learning and data engineering through the transformer era to today's foundation model--driven architectures.

Sanchit holds a Master's degree in Computer Science from RWTH Aachen University, one of Europe's leading technical universities, where his research focused on the intersection of artificial intelligence, data integration, and information retrieval. His academic work on ontology-based patent classification in medical engineering was published by Springer, reflecting an early interest in applying structured AI techniques to real-world domains.

He is a published researcher with work spanning AI, privacy, and smart systems. His publications include contributions to ontology-based classification systems (Springer, 2017), privacy-preserving data integration, smart building systems, and applied machine learning---reflecting a career-long commitment to building intelligent systems that work at the intersection of theory and practice.

\vspace{1em}

\noindent\textit{This book distills the lessons Sanchit has learned from years of building production AI systems at two of the world's largest technology companies, combined with deep engagement with the open-source and research communities. It is written for the practitioner who needs to ship AI systems that work---not just in a notebook, but in production, at scale, under real-world constraints.}

\vspace{2em}

\begin{center}
{\small\sffamily
\begin{tabular}{rl}
\textbf{Web} & \texttt{sanchitalekh.in} \\[0.3em]
\textbf{LinkedIn} & \texttt{linkedin.com/in/sanchitalekh} \\[0.3em]
\textbf{GitHub} & \texttt{github.com/salekh} \\
\end{tabular}
}
\end{center}
